### 2019-02-20 Springer_IntlCongress/author/references.tex

%%%%%%%%%%%%%%%%%%%%%%%% referenc.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% sample references
% %
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%
% BibTeX users please use
% \bibliographystyle{}
% \bibliography{}
%
\biblstarthook{}

\begin{thebibliography}{10}

    \bibitem{DarrenRowse} Darren Rowse: Rule of Thirds in Photography.
    \bibitem{Holland1975} John~H. Holland.: Adaptation in Natural and Artificial Systems In:Sgart Newsletter, 1975.
    \bibitem{Jiang2017} Yuxuan Jiang, Tomohiro Harada, and Ruck Thawonmas.: Procedural generation of angry birds fun levels using pattern-struct and preset-model In: IEEE Conference on Computational Intelligence and Games, 2017.
    \bibitem{Kaidan2015} Misaki Kaidan, Chun~Yin Chu, Tomohiro Harada, and Ruck Thawonmas.: Procedural generation of angry birds levels that adapt to the player's skills using genetic algorithm}. In: 2015 IEEE 4th Global Conference on Consumer Electronics (GCCE), pages 535--536. IEEE, oct 2015.
    \bibitem{Renz} Jochen Renz, Xiaoyu Ge, Rohan Verma, and Peng Zhang.: Angry Birds as a Challenge for Artificial Intelligence. In: Technical report.
    \bibitem{RovioEntertainmentCorporation2009} Rovio Entertainment Corporation}. In: Angry Birds, 2009.
    \bibitem{Smith2009} Gillian Smith, Mike Treanor, Jim Whitehead, and Michael Mateas.: Rhythm-Based Level Generation for 2D Platformers. In:Technical report, 2009.
    \bibitem{Standish2003}Russell~K Standish.: OPEN-ENDED ARTIFICIAL EVOLUTION. In:Technical Report~2, 2003.
    \bibitem{Sural} S.~Sural and P.K. Das.: A genetic algorithm for feature selection in a neuro-fuzzy OCR system. In: Proceedings of Sixth International Conference on Document Analysis and Recognition}, pages 987--991. IEEE Comput. Soc.
    \bibitem{Taylor2016} Tim Taylor, Mark Bedau, Alastair Channon, David Ackley, Wolfgang Banzhaf, Guillaume Beslon, Emily Dolson, Tom Froese, Simon Hickinbotham, Takashi Ikegami, Barry McMullin, Norman Packard, Steen Rasmussen, Nathaniel Virgo, Eran Agmon, Edward Clark, Simon McGregor, Charles Ofria, Glen Ropella, Lee Spector, Kenneth~O. Stanley, Adam Stanton, Christopher Timperley, Anya Vostinar, and Michael Wiser.: Open-Ended Evolution: Perspectives from the OEE Workshop in York. In: Artificial Life, 22(3):408--423, aug 2016.
    \bibitem{Yannakakis2017} Georgios~N Yannakakis and Julian Togelius.: Artificial Intelligence and Games (First Public Draft). page 359, 2017.

\end{thebibliography}