### 2019-05-14 IEEE_ConferenceOnGames/article_english.tex
hard wrap

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote.
% If that is unneeded, please comment it out.
\usepackage[style=ieee,backend=biber,style=numeric,sorting=none]{biblatex}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\addbibresource{References/article_references.bib}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Procedural generation of levels fot he angry birds videogame using
evolutionary computation\\
}

\author{\IEEEauthorblockN{ Jaime Salinas-Hernández}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Tijuana Institute of Technology}\\
Tijuana, México \\
salinas.jaime1217@gmail.com}
\and
\IEEEauthorblockN{ Mario Garcia-Valdez} \IEEEauthorblockA{\textit{Department of
Computer Science} \\
\textit{Tijuana Institute of Technology}\\
Tijuana, México \\
mariosky@gmail.com}
\and
\IEEEauthorblockN{ Given Name Surname} \IEEEauthorblockA{\textit{dept. name of
organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}}

\maketitle

\begin{abstract}
This project consisted in the generation of an evolutionary computation-based
system capable of generate and evolve the structures of which one level of the
Angry Birds game is composed, these structures are evaluated according to the
stability of the structure as well as for the complexity of said structure.

For this a level generation system was designed based on the data obtained from
the game, said data consist in the number and type of pieces that appear and the
applied gravity of the game, the individuals of the group are evaluated by a
simulation of the generated level and then checking how the structures are
affected by the gravity of the game.

The evolutionary computation system as the main objective of generating
structures based on the existent pieces of the game and evolving said pieces by
combining them in a process that simulates the rules of the Open-Ended Evolution
algorithm in which the evolution of this compounds is not inclined towards a
numeric objective rather than to extend the diversity from which the pieces may
be selected for a level.
\end{abstract}

\begin{IEEEkeywords}
genetic algorithm, procedural generated content, open-ended evolution,
evolutionary computa
\end{IEEEkeywords}

\section{Introduction}
Procedural content generation is an interesting subject on the entertainment
area, primarily in the video game industry because of the easines and quickness
provided to generate content for diferent video game projects, in this project a
evolutionary computation-based system is proposed to generate complex content
for videogames based on an evolutionary algorithm and open-ended- evolution.

For this project a procedural generation content system is proposed for the game
angry birds, angry birds is a game developed by Rovio Entrtainment in which the
objective is to destroy structures placed on the map as well as eliminating
green pig-like entities in order to obtain the highest score possible, the game
uses a gravity and collition system that allows the structures to fall and be
destroyed by means of damage as well as the use of three diferent construction
materials \textit{ice}, \textit{wood} and \textit{stone} each one with different
resistances ranging from fragile, normal and sturdy in that order, the game
mechanic is to use a slingshot to shoot a certain number of birds with different
abilities to destroy the structures and the objectives, it is possible to obtain
a high score by doing the most destruction as possible with the least birds as
you can. \cite{RovioEntertainmentCorporation2009}

While it is possible to create a evolutionary agent to play the levels and
obtain the best score the main objective of this project is to generate the
levels that can be played by naive agents or physical players, the levels
themselves need to be complex in the sense that interesting structures can be
created but it has to be possible to complete in some way
\cite{Stephenson,Stephenson2018}.

In Section 2, a background of the subjects used on this project will be
presented, this subjects, while not all being used on by different authors on
their works provide a great help on the development of the proposed system,
Section 3, contains a quick overview of some aspects used by different authors
in order to tackle this problem, their design and basic functionality, in
Section 4 the description of the current state of the project is explained, from
the basis of the development to the diferent menthods that have been used to
progress on the different sections of the project, in Section 5 we explain the
future work of the project, where we are in the development phase and how the
future implementations are to further improve the project and finally, in
Section 6 we present our conclusions

\section{Background}

In this Section we provide a quick overview of the literature about the areas
that provide the basis to the development of the current project, this section
is divided in three parts, the fisrt one gives us an explanation of what
procedural content generation entails, the second one provides an explanation on
genetic algorithms and how can they be used to solve optimization problems, and
finally, the third one provides us with an overview of Open-Ended Evolution and
what it means to have an Open-Ended Evolutionary algorithm.

\subsection{Procedural content generation}

Procedural content generation (PCG) denotes the way to automatically generate
content by algorithms instead of generating the same content by manual means
such as with different personnel.

In the videogame industry many developers use a PCG style software that produces
raw images and designs used by other personnel to improve it and use it in order
to reduce the development cycles, and the productions cost. PCG as six focus
areas listed below

PCG as six focus areas listed as follows:

\begin{itemize}
    \item Level Design: This aspect involves only the design of specific
    sections of a game, primarily those in which the player is able to interact.
    \item Audio: This aspect is focused on designing soundtracks that are able
    to react to a player specific action.
    \item Visuals: This area is focused on creating or improving visual
    representations of objects on games such as giving raw designs of players
    faces of improving the images of other graphics.
    \item Narrative: This aspect is focused on the generation of readable and
    coherent stories or plotlines of games.
    \item Gameplay: This aspect uses agents to test and evaluate the mechanic on
    which a game is based by trying to play as a human would in order to improve
    said mechanics.
    \item Game Design: This aspect is focused on generating the rulesets and mechanics of a new game. \cite{Yannakakis2017,Smith2014,shaker2016procedural}
\end{itemize}

\subsection{Genetic algorithm}
Is a method developed by John Holland on 1975 \cite{Holland1975} the genetic
algorithm (GA) is an optimization and search technique based on the principles
of genetic and natural selection, it allows a population composed of many
individuals, which by themselves represent a possible solution to the problem,
to evolve under specific selection rules in order to obtain the maximum fitness.

In other words a genetic algorithm is a method to solve optimization problems by
creating a population that will represent a possible solution and repeatedly
doing genetic operations as selection, crossover and mutation on each member of
the population in order to generate the members of the population for the next
generation that would be closer to the best solution of the problem than their
parents, the cycles in this method where the operations are made are called
generations and by moving over this the population “evolves” o in other words,
they get closer to the best possible solution  to a specific problem.

\begin{figure}[htbp]
\centerline{\includegraphics[width=90mm]{Images/ga_life_cycle.png}}
\caption{Life cycle of a genetic algorithm.}
\label{algorithm_model}
\end{figure}

The main life cycle of a genetic algorithm shown of figure \ref{algorithm_model}
is described as follows:
\begin{enumerate}
    \item Initialize a random n population of values suitable for the solution.
    \item Evaluate the fitness of each individual (solution) in the population.
    \item Check if the stopping criteria has been met.
    \item If not.
    \begin{enumerate}
        \item Introduce new individuals to the population by crossover
        operations.
        \item Mutate a random value of the new individuals.
        \item Select new cross over individuals according to their fitness for
        the next generation.
    \end{enumerate}
    \item Loop from step 2 until an end condition is satisfied.
    \item Select the best individual of the population as the solution for the
    problem
\end{enumerate}

\subsection{Open-Ended Evolution}\label{AA}
Open-ended evolution (OEE) is a variant of genetic evolution algorithm in which
the objective differs from the regular genetic algorithm one, in this case the
final objective is not to reach a final state of the evolution but to continue
evolving and creating new types of entities more complex on each new generation,
these same entities have the possibility to evolve not only inside the frame of
their type but create entirely new groups. \cite{Standish2003}

Another explanation provided by T. Taylor et al. \cite{Taylor2016,Taylor}
defines OEE has a system capable of creating novelty beyond a point where
nothing changes on the current group of interest rather than just converging to
a stable or quasi-stable state.

For this project OEE is defined simply has a process of genetic evolution that
encapsulates a group of similar processes and that is able to increase their
number by the more complex the entities become.

\section{Overview of existing work}

In this Section we present some of the different approaches to the procedural
generation of content for the angry birds video game.

M. Kaidan et al. \cite{Kaidan2015} propose a model in which the generation of
content is controlled by the skill of a player while actively playing the game,
the moment the player finishes a certain level the algorithm obtains the score
and generates the next level to be played.

J. Renz et al. \cite{Renz} created an article in which the angry birds level
generation contest mechanics of the IEEE Conference on Computational
Intelligence and Games are based on, it explains some of the solving methods
proposed by previous competitors.

Y. Jiang et al. \cite{Jiang2017} proposed the use of predefined letter style
patterns and combined the to create a small pool of word used to be combined in
order to create levels with text on them, the participants would play the game
and after that they would be asked if they liked the layout of the levels

\section{Current approach}

In this Section we describe the approach that was used for this project, The
game contains a total of 11 pieces as shown on figure \ref{piece_list} that
cannot be modified, it is not possible to add more or modify the existing ones,
with this in mind the purpose is to create bundles of this same pieces in
different locations and angles in order to create structures that can be used as
building blocks for more complex levels, in order to do this a sequence of
events needs to be done as follows:

\begin{enumerate}
\item Generate the first generation of the population
\item Run a Simulation
\item Evaluate each member of the population
\item If a member of the population has remaining pieces obtain the location and
angle of the pieces, create a bundle and add it to the pool
\item Select the parents for the next generation
\item Do the crossover and mutation operations
\item Repeat the cycle
\end{enumerate}

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/list_pieces.png}}
\caption{List of pieces in the angry birds game.}
\label{piece_list}
\end{figure}

The rules for the competition specify that a configuration file must be used to
generate the levels of the game in a limited amount of time, the infornmation on
the configuration file is first the number of \textbf{\textit{diferent}} levels
that must be generated, the second value is the diferent materials \textit{not
allowed} to be used, for instance a group of generated levels must not have
circles which composition material is ice or stone moreover it des not have to
integrate a triangle piece with ice, stone or wood material so this combinations
must be removed from the pool of options, the third configuration value is the
\textit{number of pigs} that must be placed in a level, and finally the
\textit{limit} of time that the system has to create the required amount of
levels.

After taking into account this restrictions the strcture of our current approach
had to be changed, so this section will be divided in tow subsections, the first
will explain the approach that was used and after that second will explain the
modifications that had to be made in order to integrate the configuration files
for the competition.

\subsection{Original approach}

Since the algorithm has to generate more complex structures the remaining pieces
of a member of the population have to be used has a starting point, this new
additions to the piece pool are added to the population by the mutation
operation, the objective in the first phase of the project is to generate
structures that are capable of maintaining their own balance as shown on figure
\ref{test_old}, since the objective of open-ended evolution for this project is
to add new bundles to the pool of options the algorithm will first run a
simulation with a group of pieces and after it the ones that were not destroyed
by falling because of gravity will be saved and added to the pool of options.

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/simulation_bef_aft_example.png}}
\caption{Structure at beginning and end of a simulation.}
\label{test_old}
\end{figure}

In order to add a bundle of pieces as a new component first it has to be
measured, an imaginary bounding box is defined by calculating the lengths of all
the pieces, finding the borders of said pieces and find the points that could
define a box around all of them, calculate the height and width then calculate
the center point of said box and calculate the offsets from the center of the
group to the center of each respective piece as shown on figure
\ref{bounding_boc_calc}, this information is needed in order to add it as a new
element so it can be used to place the pieces in their respective new positions
when the files needed for the simulation are being created.

The way to represent the elements of a member of the population is by using the
id of each piece and the new added ones has a pointer to the respective data
that needs to be obtained, this way and individual can be represented as shown
on figure \ref{old_chrom} where each chromosome represents a certain item in the
pool of pieces, the numbers for this elements are assigned in a first come first
served basis, the numbers are infinitely incremented and this same pieces can be
repeated on each individual of the population

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/bounding_box_calculation.png}}
\caption{Bounding box calculation.}
\label{bounding_boc_calc}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/chromosome_chain_example.png}}
\caption{Chromosome chain of an individual.}
\label{old_chrom}
\end{figure}

Once the pieces have been added and the individuals are ready to be evaluated a
simulation on the game environment is scheduled by order in the population, all
the individuals have a maximum of ten seconds to maintain balance or reach
absolute stability, if absolute stability is reached it means one of two things,
first the generated structure is completely stable or second, the tower fell to
either side and the pieces were destroyed or stopped moving by gravitational
pull, when either of this happens by more than two seconds the simulation ends
regardless of time remaining and the next individual is simulated.

The evaluation process for this project is still in the development stages,
however in order to check that the evolution is working has it should a criteria
has been temporarily made as follows, each individual in the population has 100
points so before the simulation starts each of them are viable candidates to be
a elite member of the population regardless relative height or complexity, then
the remaining of the process is separated in three phases has follows.

First, after each of the individual simulations are over the evaluation process
takes place, in this phase the resulting level is checked for pieces, each one
of them is counted and then the resulting number is compared to that of the
initial quantity of pieces that were originally put in before the simulation in
order to calculate the first element of the evaluation process, each piece
represents a certain percentage of 100\% so the missing pieces are subtracted
from the total.

Second, after removing percentage according to the missing pieces an individual
remaining pieces are checked for their positions in comparison to the original
ones before the simulation, using the formula 1 each piece within the individual
is evaluated according to the center point of the piece, if the piece moved more
than a certain threshold half the percentage for one piece is removed from the
remaining percentage, after this using the formula 2 the position of the piece
is further checked this time by the difference in angle from the original one,
in case the difference of angles is more than a certain threshold half a piece
value is further removed from the total.

\begin{equation}
    \begin{aligned}
    error_{xy} = \left\{ \,
        \begin{IEEEeqnarraybox}[][c]{l?s}
            \IEEEstrut
            0.1 & if $ 0.08 > d $, \\
            \frac{100}{lenght\_pieces} * 0.5 & if $ 0.08 < d $.
            \IEEEstrut
        \end{IEEEeqnarraybox}
    \right. \\
    where: d=\sqrt{(x_2 - x_1)^2 + (y_2 - y_1)^2}    
    \end{aligned}
    \label{error_xy}
\end{equation}

\begin{equation}
    \begin{aligned}
    error_{r} = \left\{ \,
        \begin{IEEEeqnarraybox}[][c]{l?s}
            \IEEEstrut
            0 & if $ -5 \le r \le 5 $, \\
            \frac{100}{lenght\_pieces} * 0.5 & if $ r < 5 or r > 5 $.
            \IEEEstrut
        \end{IEEEeqnarraybox}
    \right. \\
    where: r= |rotation_0| - |rotation_f|   
    \end{aligned}
    \label{error_angle}
\end{equation}

Finally, the total height of the remaining group of pieces is calculated by
combining the bonding boxes of all pieces, then the individuals are ordered by
the obtained percentage and then further ordered by the total height, this way
the most balanced ones will have the best probability to be selected together
for a crossover operation to create the next generation.

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/result_example.png}}
\caption{Result samples of current system.}
\label{results_old}
\end{figure}

Since there is a need to create complex structures than can cover a great
section of the play map, a method to evaluate the distribution is needed, for
this instance we propose using the rule of thirds \cite{DarrenRowse} which is a
guideline used while creating imagen in which the purpose is to create
aesthetical pleasant images where the focus point or the most important area is
at the center of the image as shown on figure \ref{rule_of_thirds}, using this
rule different distribution groups or masks are created has shown on figure
\ref{rule_of_thirds_applied} in order to use them before the simulation of an
individual takes place in order to distribute the pieces in a way that can cover
more area and create more balanced levels instead of using a tower like
distribution as shown on figure \ref{results_old}.

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/ruleofthirds_example.png}}
    \caption{Rule of thirds example.}
    \label{rule_of_thirds}
\end{figure}

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/mask_distribution.png}}
    \caption{Rule of thirds applied to an individual.}
    \label{rule_of_thirds_masks}
\end{figure}

Using this same distribution a new evaluation will be added in which according
to how much of the area is filled a score will be given and this score will be
used as a new sorting value at the moment of selecting individuals for the
crossover operation, since as shown on figure \ref{rule_of_thirds_masks} there
is a possibility that a huge mask is given to an individual it can be possible
that the amount of pieces will not be able to cover the full area given, so in
order to prevent this kind of problems we also proposed to give each mask a
value that refers to the minimum amount of pieces and height required to fill
the most of the mask then check the amount of pieces and height before
simulation for a single individual and then randomize which mask of the ones
that can be used will be assigned to that individual, this way the results shown
on figure \ref{rule_of_thirds_result} are obtained, this however gave us a new
problem because as shown in the image and the previous image as well the were
instances in which a certain combination of pieces could be to big for the area
and this pieces resulted in a level that had structural flaws from the beginning
which cold provoke the entire structure and immeadiately nearby others to fall
at the beginning of the simulation however it was able to

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/chromosome_thirds.png}}
    \caption{Rule of thirds applied to an individual.}
    \label{rule_of_thirds_applied}
\end{figure}

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/result_example_thirds.png}}
    \caption{Results obtained with rule of thirds.}
    \label{rule_of_thirds_result}
\end{figure}

\subsection{New approach}

The first differece that the new approach as is that the masks that are assigned
to each individual are no longer pre-generated ones, this is because we decided
that pre-generating them would not allow the generator to \textit{evolve}
because the individuals would be limited to a certain number of possibles
combinations which will in turn prevent diversity in the population, so in order
to prevent this the modification to this section allows the individuals to
generate a mask in which a number of columns is defines by default but the
amount of pieces in each one is random, however the total of assigned pieces in
all columns must be the same as the total pieces alloweed in each individual
which is a number assigned at before beginning the system, an example of this
can be seen in figure \ref{new_chrom} where the total of pieces in a individual
is 18 and this pieces are distributed on all the columns available. This is
further explained in this section.

The competition also requires that a generator gives a set output a file is used
to simulate the ruleset provided at a competition, this \textit{text file
(parameters.txt)} provides the specifications that must be met in the
competition, the parameters in the file are as follows:

\begin{itemize}
    \item Total required levels
    \item Restrictions to pieces or piece-material combinations that cannot be
    placed or used in the levels, in this case if the file specifies that a
    \textit{CIRCLE} made out of \textit{ICE} cannotplaced the system must be
    able to verify that said combination is not used.
    \item Total number of pigs, this value can be a single value or a range of
    values that represent how many pigs must be placed on a level.
    \item The amount of time in which the system must give an output.
\end{itemize}

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/chromosome_chain_new_model.png}}
\caption{New chromosome chain.}
\label{new_chrom}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/new_model.png}}
\caption{New system diagram based on the parameter file and the new chromosome representation.}
\label{new_model}
\end{figure}

Using this as a base a new system diagram is proposed as shown of figure
\ref{new_model} in which the file with parameters is integrated with the input
data  the system requires, the way the individuals of the population are further
presresented and the way the simulation phase is taken, the method to
representate the individuals is proposed by dividing the individual on three
layers represented as follows:

\begin{itemize}
    \item Element layer: also known as pieces layer, this layer represent the
    elements assigned to a particular individual by an array of numbers from 0
    to infinite where each number represents a reference to the class of a piece
    in the selection pool including the ones generated by the OOE algorithm, the
    representetion of this is was shown in figure \ref{old_chrom}.
    \item  Mask layer: this layer is represented by a 7 digit array in which the
    value assigned in each column represents the amount of pieces from the
    element layer that will be assigned to said column, a representation of this
    is shown on figure \ref{mask_layer}.
    \item Enemy layer: also known as pig layer, this layer is represented by two
    elments a single numeric value representing the total amount of pigs
    assigned to the level and an array of locations for the placement of this
    same pigs, a representation of it is shown on figure \ref{enemy_layer}. 
\end{itemize}

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/mask_layer.png}}
    \caption{Mask array with column and fenotype representation.}
    \label{mask_layer}
\end{figure}

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/enemy_layer.png}}
    \caption{Amount of pigs in a level, their location relative to the mask layer (green) and fenotype representation.}
    \label{enemy_layer}
\end{figure}

In order to select the parents of the next generation the parents are ordered by
their fitness, then the system can choose between 2 options for the selection,
tournament selection or roulette selection:

\begin{itemize}
    \item Tournament selection: in order to select the parent by this method the
    system checks the amount of crossovers assigned, then calculates the ammount
    of parent required to compete in pairs, then by doing a descendent selection
    the slots for the tournament are filled and the winners are calculated in
    each group.
    \item Roulette selection: this method of selection was first decided to be
    used directly by giving a proportional amount of chances to be selected
    according to the fitness of a individual, however since one of the
    objectives is to obtain variety of levels the roulette must first be ranked
    in order to ensure that low fitness individuals can be selected as parents.
\end{itemize}

As for the crossover operations the system currently uses two types, single
point crossover and double point crossover, however this crossover operation is
not only applied to the piece array of an individual, but it is also applied to
the mask of said individual, this way the mask can ensure that some pieces of
some childs in the population are now placed in the level, this can bring the
possibility that a mask that uses less pieces that it normally as can obtain a
better fitness than one that uses all of them. An example of this two crossover
thypes cna be seen in figure \ref{crossover} in which the mask of an individual
is combined with another and the resulting mask for the childs have different
amount of pieces to be placed.

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/crossover.png}}
    \caption{Single point crossover(up) and double point crossover(bottom) aplied to an individual mask.}
    \label{crossover}
\end{figure}

The mutation operations of the generator work with the pieces layer of an
individual, the four types of mutations used are:

\begin{itemize}
    \item Add or remove elements from the piece list
    \item Change the piece that is being used in a point of the list
    \item Change the material a piece is made offsets
    \item Change the x and y coordinates of a piece
\end{itemize}

This mutations do not have to be used all at once on a single individual but
depending on the probabilities they can happen all at once, since most of the
pieces on a individual are mostly the same type of material then at least the
material mutation can occurr on all or most of the individuals to ensure that
all levels will not have all the pieces with the same material which could
render a level \textit{"boring"} by the fitness function.

Since this group of mutations further ensures diversity in the populations a
different way to evaluate the individuals is needed, by using the previous
proposed method as a base a way to evaluate the \textit{"diversity"} and
\textit{"uniqueness"} of a individual, for this we first remove the \textit{x}
and \textit{y} position as well as the \textit{angle} errors in order to add two
new evaluation formulas as follows:

\begin{itemize}
    \item Hamming distance: a simle explanation for the \textit{Hamming
    distance} is that it is a calculation applied to a pair of strings where the
    obecive is to find the minimum required number of substitutions required in
    order to convert a string exactly the same as the other, its formula is
    shown in equation \ref{hamming_distance}, since the formula itself locates
    the minimum required number of substitutions then we compare the individual
    with the rest of the population trying to found the maximum number of
    substitutions required efectively using the formula to give a better fitness
    to those individuals that are different from the group.
    \item Shannon entropy: the shannon entropy is used on information theory to
    calculate the disorder or uncertainty, in this case it will be used to
    determine if a level is \textit{"boring"} or \textit{"interesting"} based on
    the amount of entropy calculated according to the pieces used in a level
    where a low entropy is boring because if a determined level as a lot of
    repeated pieces then it will look simple, the formula used is shown in
    equation \ref{shannon_entropy} where the elements calculated are the pieces
    of a level and their appearance probability against all others.
\end{itemize}

\begin{equation}
    \begin{aligned}
    d = min \left\{ \ d(x,y): x,y \in C, if \: x \neq y, 1 \: else \: 0 \: \right\} \
    \end{aligned}
    \label{hamming_distance}
\end{equation}

\begin{equation}
    \begin{aligned}
    S = - \Sigma_i P_i \log P_i
    \end{aligned}
    \label{shannon_entropy}
\end{equation}

Using all this configurations the system is capable of creating levels, some of
them playable as shown in figure \ref{result_1} however the levels that can be
generated are far from optimized to be played on, this is an issue that must be
changed in future instances but the generator itself is able to be used as a
base for a more complex one, and as shown in the future work section this issues
can be solved by different means.

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/Result_n2.png}}
    \caption{Example of resulting level with current system.}
    \label{result_1}
\end{figure}

\section{Future work}

The system as been able to generate levels that fulfill the requirements
 provided by the parameters file, however a way to evaluate that the levels
 generated are indeed \textit{"fun"} or \textit{"interesting"} is still needed,
 at this point we have two was to evaluate the results:
 \begin{itemize}
    \item User survey, in case a survey is used as an evaluation tool then first
    the system has to generate the levels and a group of people must participate
    by playing the levels and answering a set of questions afterwards, the
    questions themselves must try to capture the feeling of a person towards the
    game like questioning if a level was too easy or too dificult to solve, if
    the generated structures look interesting or if the level required them to
    strategize how to use the provided birds to solve the level, the survey must
    be applied to a equal quantity of people of three diferent game skill
    levels,\textit{"non video game players"}, \textit{"casual video game
    players"} and \textit{"avid video game players"}, this way we can evaluate a
    level or the entire system by knowing if all skill levels did not enjoyed a
    level because it was too simple or easy.
    \item By using a naive agent, if a naive agent is used to play the levels we
    can obtain data of how the agent solves the levels placed before them, and
    the way to evaluate if the level is or not \textit{"interesting"} will be by
    checking how easily the aggent is able to solve a level, the point here is
    to have a level that cannot be solved by a single movement but to have one
    that a human player has to think a while to solve tryingto place the
    difficulty of a level just at the middle point between easy and hard
    difficulty levels.
\end{itemize}

Other aspects that the system need to improve is in the processing time
required, since the current way to calculate the fitness and create the next
generations of a population the system requires the system to execute
simulations of the levels on every generation that passes therefore increasing
the time it takes to complete an entire cycle of execution, so in order to
change this we require to create a way to evaluate a level using only the
information of the pieces, locations and pigs locations and using this data to
predict the way a level will behave when placed in a real simulation environment
therefore drastically reducing execution times.

Finally a different method to evaluate the individuals must be created, this is
because the current evaluation method allows us to generate playable levels,
this levels are far from ideal in the way that they only use a certain range of
items present in the game therefore reducing the potential of a level, we expect
to make the current fitness function work the most eficient that it can but
another possible solution is found a good combination of formulas to generate
the fitness function.

\section{Conclusions}

The use of artificial intelligence in videogames is a common thing almost since
before the first generation of videogames on 1972 as simply entities that would
follow a set group of actions but as the generation kept advancing the need for
more intelligent entities appeared, as of this generation some videogames use AI
as ways to improve the interaction between man and game, but also using it in
order to facilitate the way videogames are created, so for this project the
focus of using evolutionary algorithms as great interest, as a starting point a
rather simple game can be used as a workbench of the work but it also helps us
define how much more can be done when all the required objects are obtained,
improve on the same system and find ways for it to be used on different other
games.

\printbibliography
\newrefcontext[sorting=ydnt]

\end{document}